Designing a modular architecture for building web-based psychological experiments

### 2021-08-04 paper.tex
Use thesis template for styling

\documentclass[a4paper,10pt]{paper}
\usepackage[margin=1.25in,footskip=0.25in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{csvsimple}
\usepackage{hyperref}

@@

\usepackage{listings}
\usepackage{color}
\lstset{
   language=html,
   numbers=left,
   numberstyle=\tiny,
   breaklines=true,
   showstringspaces=false,
   basicstyle=\ttfamily,
   keywordstyle=\color[rgb]{0,0,1},
   commentstyle=\color[rgb]{0.133,0.545,0.133},
   stringstyle=\color[rgb]{0.627,0.126,0.941}
}

@@

@@

\maketitle
\begin{abstract}
In the social and behavioral sciences, web-based experiments allow experimentors to recruit larger, more diverse subject pools with higher statistical power, and can speed up experiments dramatically, while being portable to any device with a browser. While web-based experiments have become increasingly popular, methods to construct them have been lacking in flexibility and simplicity. We introduce magpie, a modern, component-based, declarative JavaScript framework based on Vue.js for building web-based experiments of varying complexity with ease, including the possibility to have participants interact virtually.
\end{abstract}
\newpage

@@

\section{Introduction}
Web-based experiments are gaining more popularity in behavioural research, because of their flexibility and efficiency. From our analysis, we find that many existing tools for realizing such experiments lack sufficient flexibility and expressiveness. This lead to the creation of magpie, both a server for managing experiments and a client-side framework for implementing them.

@@

\section{Requirements}
In our analysis of prior work as well as in the design of magpie, we focus on the following features we deem desirable.

@@

\section{Analysis of prior work}
There are a variety of pre-existing software packages attempting to aid in the construction of web-based experiments. In the following we examine some prominent candidates.

@@

\section{Design}

\subsection{Design goals and choices}
A main goal of designing magpie was to provide implementers with a conceptually simple framework for building experiments that should be employable as a tool for teaching the basics of experimental design. At the same time the framework should be general and expressive enough to facilitate building complex interactive experiments without major software engineering skills.

@@

\subsection{Overview}
Magpie consists of two parts that each have their own responsibilities: a server-side service and a client-side framework. These two parts, although integrating with each other, are largely independent. The server handles experiment management, configuration and data storage. It is designed to be the backend for multiple different experiments that are possibly hosted elsewhere.

@@

\subsubsection{Conceptual Framework}

@@

\subsection{Components}
Magpie provides basic stimulus components like playing audio files, video files, displaying image files, text, a self-paced reading component, a tone Synthesizer and a canvas for displaying dynamic visual stimuli. Responses can be recorded using dropdown inputs, a forced choice input, a single and a matrix multiple choice input, a slider input, a rating scale, a free text box, a key press input, a constant sum range slider, an image selection input, a text completion input and a rank order input.

@@

\subsection{Dynamic experiments}
Iterated experiments allow using data from past participants in later trials for new participants. This allows the creation of, for example, iterated narration experiments (similar to the game of telephone).
Magpie requires three basic settings on the server for running an iterated experiment: The number of variants, which defines how many different experimental settings there will be; the number of chains, which defines how many closed participant groups there will be; and the number of generations, which defines how many participants are in one chain.

\subsection{Interactive experiments}
Interactive experiments allow participants to interact with each other in some form, for example as rational decision-makers with strategic interactions in a mathematical model, as is frequently described in Game theory. On the client-side, magpie provides components for structuring interactive experiments, such as a screen to wait until the required number of participants in the group are online. To allow the flexible realization of any kind of interactivity, magpie allows sending and listening to events between individual participant sessions. Additionally, it keeps track of which participants currently see the same screen and assigns pseudonymous names and colors to all participants that can be used in experiments for easier identification during interaction.

\subsection{Technical realization}
Technologically, the client-side framework is based on Vue.js \citep{You2014}, an open-source front-end JavaScript framework for building web-based user interfaces and single-page applications. It facilitates declarative, reactive rendering and component composition, which is in contrast to traditional imperative programming, and greatly eases reasoning about and implementation of applications. Declarative rendering is facilitated by allowing to extend normal HTML code with behavior-invoking directives and stateful custom elements.

@@

\subsubsection{Example 1: Mental rotation experiment}
The following example shows the code for a simple mental rotation experiment. The initial screen would display instructions to the participant, but is excluded here for brevity. Afterwards, the code iterates over a list of trial objects, defined elsewhere and constructs a new KeypressScreen for each. KeypressScreen is a built-in screen with four life cycle phases: Two seconds pause and one second fixation, as defined in the screen's properties. Third and fourth phase are merged together, by default, where the third phase displays the contents of the \texttt{\#stimulus} slot and the fourth phase allows responding using the F or the J key on the keyboard, which will indicate whether the participant thinks the two models in the picture are the same or different. The screen automatically records response time and response. (Find more information on built-in screens can be found at \url{https://magpie-manual.netlify.app/01_designing_experiments/00_built-in_screens/})

@@

\subsubsection{Example 2: Simon Task}
The following example shows the code for a Simon Task experiment. The code in this example leaves out the instructions to the participants. The code iterates over a list of trial objects, defined elsewhere and constructs a new KeypressScreen for each.

@@

\subsubsection{Example 3: Public goods experiment (interactive)}
The following example shows the code for a public goods experiment, an interactive paradigm from behavioral economics. The code in this example leaves out the instructions to the participants. Before the trials start, the ConnectInteractiveScreen makes sure the socket connection to the server is active and the necessary number of participants have joined the experiment. The code then iterates ten times, and constructs a new raw Screen for each iteration.

@@

\section{Limitations and future directions}
A limitation of using Vue.js is the required build step for publishing experiments. Thus, making changes to existing experiments is only possible when the required toolchain is installed. This is a trade-off towards the usability of the live-reloading and modularity features of modern JavaScript frameworks. Magpie also still lacks built-in screens for a range of trial types, such as random dot kinematograms or visual search tasks. A key limitation is the need to write code in order to build experiments with magpie, a trade-off we find worthwhile to allow implementers greater flexibility. However, this does not exclude the possibility of building a graphical editor in the future.

@@

Another avenue for improvement is interoperability with JATOS \citep{Lange2015} and Sona \citep{SonaSystems2002}.
\section{Conclusion}
Web-based scientific experiments offer higher statistical power at a lower cost and effort with only marginal disadvantages. Magpie's client-side leverages the advantages of modern JavaScript frameworks to offer a simple yet powerful framework for building such experiments. Its conceptual structure facilitates realizing basic as well as complex and even interactive experimental paradigms, while striking a balance between usability and expressiveness. Provided under an open source license, magpie facilitates open science, and eases the replication process. Fully relying on the power of a Turing-complete programming langauge for implementing experiments allows extending the basic framework with additions such as Optimal Experiment Design methods with ease.

>>>>

\documentclass[a4paper,11pt]{scrreprt}
\usepackage[margin=1.25in,footskip=0.25in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{csvsimple}
\usepackage{hyperref}
\usepackage{authoraftertitle}
\usepackage{setspace}

@@

% Line for signature.
\newcommand{\namesigdate}[1][5cm]{%
	\vspace{5cm}
	{\setlength{\parindent}{0cm}
	\begin{minipage}{0.3\textwidth}
		\hrule 
		\vspace{0.5cm}
		{\small city, date}
	\end{minipage}
	 \hfill
	\begin{minipage}{0.3\textwidth}
		\hrule
		\vspace{0.5cm}
	    {\small signature}
	\end{minipage}
	}
}

\usepackage{listings}
\usepackage{color}
\lstset{
   language=html,
   numbers=left,
   numberstyle=\tiny,
   breaklines=true,
   showstringspaces=false,
   basicstyle=\ttfamily\singlespacing,
   keywordstyle=\color[rgb]{0,0,1},
   commentstyle=\color[rgb]{0.133,0.545,0.133},
   stringstyle=\color[rgb]{0.627,0.126,0.941},
}

@@

\linespread{1.5} % set line spacing

@@

\begin{titlepage}
	\begin{flushleft}
		Universität Osnabrück\\
		Fachbereich Humanwissenschaften\\
		Institute of Cognitive Science
	\end{flushleft}

	\vspace{2cm}
	\centering{
		Bachelor thesis\vspace{1cm}\\
		\textbf{\Large{\MyTitle}}
		\vspace{1cm}\\
		\begin{tabular}{c}
			\MyAuthor                          \\
			977686                             \\
			Bachelor's Program Cognitive Science
		\end{tabular}}
	\vspace{1cm}

	\begin{tabular}{ll}
		First supervisor:  & Prof. Dr. Michael Franke\\
		                   & Institute of Cognitive Science                   \\
		                   & Osnabrück                       \\\\
		Second supervisor: & Britta Grusdt, M.Sc.         \\
		                   & Institute of Cognitive Science \\
		                   & Osnabrück
	\end{tabular}

\end{titlepage}

\chapter*{Declaration of Authorship}
I hereby certify that the work presented here is, to the best of my knowledge and belief, original and the result of my own investigations, except as acknowledged, and has not been submitted, either in part or whole, for a degree at this or any other university.

\namesigdate
\pagebreak

\begin{abstract}
In the social and behavioral sciences, web-based experiments allow experimentors to recruit larger, more diverse subject pools with higher statistical power, and can speed up experiments dramatically, while being portable to any device with a browser. While web-based experiments have become increasingly popular, methods to construct them have been lacking in flexibility and simplicity. We introduce magpie, a modern, component-based, declarative JavaScript framework based on Vue.js for building web-based experiments of varying complexity with ease, including the possibility to have participants interact virtually.
\end{abstract}
\newpage

@@

\chapter{Introduction}
Web-based experiments are gaining more popularity in behavioural research, because of their flexibility and efficiency. From our analysis, we find that many existing tools for realizing such experiments lack sufficient flexibility and expressiveness. This lead to the creation of magpie, both a server for managing experiments and a client-side framework for implementing them.

@@

\chapter{Requirements}
In our analysis of prior work as well as in the design of magpie, we focus on the following features we deem desirable.

@@

\chapter{Analysis of prior work}
There are a variety of pre-existing software packages attempting to aid in the construction of web-based experiments. In the following we examine some prominent candidates.

@@

\chapter{Design}

\section{Design goals and choices}
A main goal of designing magpie was to provide implementers with a conceptually simple framework for building experiments that should be employable as a tool for teaching the basics of experimental design. At the same time the framework should be general and expressive enough to facilitate building complex interactive experiments without major software engineering skills.

@@

\section{Overview}
Magpie consists of two parts that each have their own responsibilities: a server-side service and a client-side framework. These two parts, although integrating with each other, are largely independent. The server handles experiment management, configuration and data storage. It is designed to be the backend for multiple different experiments that are possibly hosted elsewhere.

@@

\subsection{Conceptual Framework}

@@

\section{Components}
Magpie provides basic stimulus components like playing audio files, video files, displaying image files, text, a self-paced reading component, a tone Synthesizer and a canvas for displaying dynamic visual stimuli. Responses can be recorded using dropdown inputs, a forced choice input, a single and a matrix multiple choice input, a slider input, a rating scale, a free text box, a key press input, a constant sum range slider, an image selection input, a text completion input and a rank order input.

@@

\section{Dynamic experiments}
Iterated experiments allow using data from past participants in later trials for new participants. This allows the creation of, for example, iterated narration experiments (similar to the game of telephone).
Magpie requires three basic settings on the server for running an iterated experiment: The number of variants, which defines how many different experimental settings there will be; the number of chains, which defines how many closed participant groups there will be; and the number of generations, which defines how many participants are in one chain.

\section{Interactive experiments}
Interactive experiments allow participants to interact with each other in some form, for example as rational decision-makers with strategic interactions in a mathematical model, as is frequently described in Game theory. On the client-side, magpie provides components for structuring interactive experiments, such as a screen to wait until the required number of participants in the group are online. To allow the flexible realization of any kind of interactivity, magpie allows sending and listening to events between individual participant sessions. Additionally, it keeps track of which participants currently see the same screen and assigns pseudonymous names and colors to all participants that can be used in experiments for easier identification during interaction.

\section{Technical realization}
Technologically, the client-side framework is based on Vue.js \citep{You2014}, an open-source front-end JavaScript framework for building web-based user interfaces and single-page applications. It facilitates declarative, reactive rendering and component composition, which is in contrast to traditional imperative programming, and greatly eases reasoning about and implementation of applications. Declarative rendering is facilitated by allowing to extend normal HTML code with behavior-invoking directives and stateful custom elements.

@@

\subsection{Example 1: Mental rotation experiment}
The following example shows the code for a simple mental rotation experiment. The initial screen would display instructions to the participant, but is excluded here for brevity. Afterwards, the code iterates over a list of trial objects, defined elsewhere and constructs a new KeypressScreen for each. KeypressScreen is a built-in screen with four life cycle phases: Two seconds pause and one second fixation, as defined in the screen's properties. Third and fourth phase are merged together, by default, where the third phase displays the contents of the \texttt{\#stimulus} slot and the fourth phase allows responding using the F or the J key on the keyboard, which will indicate whether the participant thinks the two models in the picture are the same or different. The screen automatically records response time and response. (Find more information on built-in screens can be found at \url{https://magpie-manual.netlify.app/01_designing_experiments/00_built-in_screens/})

@@

\subsection{Example 2: Simon Task}
The following example shows the code for a Simon Task experiment. The code in this example leaves out the instructions to the participants. The code iterates over a list of trial objects, defined elsewhere and constructs a new KeypressScreen for each.

@@

\subsection{Example 3: Public goods experiment (interactive)}
The following example shows the code for a public goods experiment, an interactive paradigm from behavioral economics. The code in this example leaves out the instructions to the participants. Before the trials start, the ConnectInteractiveScreen makes sure the socket connection to the server is active and the necessary number of participants have joined the experiment. The code then iterates ten times, and constructs a new raw Screen for each iteration.

@@

\chapter{Limitations and future directions}
A limitation of using Vue.js is the required build step for publishing experiments. Thus, making changes to existing experiments is only possible when the required toolchain is installed. This is a trade-off towards the usability of the live-reloading and modularity features of modern JavaScript frameworks. Magpie also still lacks built-in screens for a range of trial types, such as random dot kinematograms or visual search tasks. A key limitation is the need to write code in order to build experiments with magpie, a trade-off we find worthwhile to allow implementers greater flexibility. However, this does not exclude the possibility of building a graphical editor in the future.

@@

Another avenue for improvement is interoperability with JATOS \citep{Lange2015}, an experiment management backend, and Sona \citep{SonaSystems2002}, a management system for university-assigned experiment participation.

\chapter{Conclusion}
Web-based scientific experiments offer higher statistical power at a lower cost and effort with only marginal disadvantages. Magpie's client-side leverages the advantages of modern JavaScript frameworks to offer a simple yet powerful framework for building such experiments. Its conceptual structure facilitates realizing basic as well as complex and even interactive experimental paradigms, while striking a balance between usability and expressiveness. Provided under an open source license, magpie facilitates open science, and eases the replication process. Fully relying on the power of a Turing-complete programming langauge for implementing experiments allows extending the basic framework with additions such as Optimal Experiment Design methods with ease.
